---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Prediction Assignment Writeup

## Background

In this project,I will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

### Data loading and processing
```{r}
setwd("C:/Users/rr111836/Desktop/Studies/Coursera/Assignment 3")

library(readr)
training<-read.csv("pml-training (1).csv",na.strings = c("NA", "#DIV/0!", ""))
testing <- read.csv("pml-testing (1).csv",na.strings = c("NA", "#DIV/0!", ""))
```

### Loading Required package

```{r}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(301)
```






### Removing columns that contains NA values and irrelevant variables and Partioning the training set into training and crossvalidation datasets

```{r}
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```
### remove variables with Nearly Zero Variance

```{r}

n0var <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -n0var]
TestSet  <- TestSet[, -n0var]
dim(TrainSet)
```
```{r}
dim(TestSet)
```

###Remove Variables that are mostly NAs

```{r}
AllNA <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```

```{r}
dim(TestSet)
```


```{r}
# remove identification only variables (columns 1 to 5)
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

###check correlation among variables



```{r}
M <- cor(TrainSet[, -54])
corrplot(M, method="circle")
```

### Random Forest method

```{r}
# plot matrix results
# model fit
set.seed(3408)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```

### Prediction

```{r}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

### Plot Matrix Results

```{r}
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```

### Decision Tree

```{r}
# model fit
set.seed(3408)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modFitDecTree)
```


### Prediction on Test dataset

```{r}
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

### Plot matrix results

```{r}
plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 4)))
```


### Applying the selected Model to the Test Data

```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```


















